Overview

### 2024-01-05 analysis/001_overview.Rmd
Algin plot design

```{r area_oa_over_the_years}
oa_ta_plot_df <- ta_df |>
  select(cr_year, ta_oa, no_ta) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  inner_join(year_all)
  # plot
oa_ta_plot <- ggplot(oa_ta_plot_df, aes(cr_year, value / n, fill = type, group = type)) +
  geom_area() +
  scale_fill_manual(NULL, values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(), limits = c(0, .15), breaks = c(0, .05, .1, .15)) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(legend.position="top", legend.justification = "left", axis.text.x = element_blank())
```

```{r ta_share_plot}
ta_share_plot <- ta_df |>
  select(cr_year, ta_oa_prop, no_ta_prop) |>
  pivot_longer(cols = c(ta_oa_prop, no_ta_prop), names_to = "type", values_to = "value") |>
  filter(!cr_year %in%  c("2017", "2023"), type == "ta_oa_prop") |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = 1)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25, 0.5),
                     labels = scales::percent) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = "Publication Year")
```

@@

summary_all <- inner_join(ta_df, jn_year_all, by = c("cr_year", "issn_l")) |>
  mutate(prop_ta_oa = ta_oa / n,
         prop_no_ta = no_ta / n) |>
  ungroup() |>
  distinct(cr_year, issn_l, prop_ta_oa, prop_no_ta) |>
  pivot_longer(cols = c(prop_ta_oa, prop_no_ta))
# plot
box_ta_prop <- summary_all |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  # With at least one OA article
  filter(value > 0) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, fill = name)) +
  geom_boxplot(outlier.shape = NA, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  coord_cartesian(ylim = c(0,0.65)) +
  scale_fill_manual(NULL, values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement"), guide = FALSE) +
  theme_minimal() +
  labs(y = NULL, x = "Publication Year")

@@

publisher_prop <- ggplot(publisher_summary_df, aes(gsub("^20", "'", as.character(cr_year)), prop, fill = type, group = type)) +
  geom_area() +
  facet_wrap(~publisher, nrow = 1) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(),
                     breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(legend.position="top", legend.justification = "left", axis.text.x = element_blank())

publisher_boxplot <- publisher_plot_df |>
  filter(value > 0) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.65)) +
  facet_wrap(~publisher, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )

publisher_ta_prop_plot <- publisher_ta_prop |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
                     labels = scales::percent) +
  facet_wrap(~publisher, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank())
```

@@

```{r subject_plots}
subject_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  select(cr_year, top_level, contains("frac"),-oa_frac) |>
  pivot_longer(
    cols = c(ta_oa_frac, no_ta_fractions),
    names_to = "type",
    values_to = "value"
  ) |>
  mutate(prop = value / articles_frac) |>
  ggplot(aes(
    gsub("^20", "'", as.character(cr_year)),
    prop,
    fill = type,
    group = type
  )) +
  geom_area() +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_fill_manual(
    "Hybrid OA",
    values = c(
      no_ta_fractions = "#b3b3b3a0",
      ta_oa_frac = "#56B4E9"
    ),
    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

subject_ta_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  mutate(prop = ta_oa_frac / oa_frac) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
                     labels = scales::percent) +
  facet_wrap(~top_level, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank())
subject_box_trend <- subject_df |>
   # with at least one
  select(-ajcs, -subject_area, -esac_publisher) |>
  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(value > 0) |>
  filter(top_level != "Multidisciplinary", !cr_year %in% c("2017", "2023")) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.55)) +
  facet_wrap(~top_level, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )
```

@@

# Prepare facet heading
facet_labels <- country_multiple_df |>
  group_by(country_name_fct) |>
  summarise(total = sum(all_total)) |>
  mutate(prop = total / sum(total)) |>
  ungroup() |>
  # Highlight Germany
  mutate(total_string = case_when(
    total < 1000 ~ as.character(total),
    total < 1000000 ~ paste0(format(round(total / 1e3, 1), trim = TRUE), "K"),
    total >= 1000000 ~ paste0(format(round(total / 1e6, 2), trim = TRUE), "M")
  )) |>
  mutate(
    facet_labels = glue::glue(
      '{country_name_fct}<br>
      {total_string} | {paste(round(prop * 100, 1), "%")}'
    )
  ) |>
  distinct(country_name_fct, facet_labels)

@@

```{r, country_plot, fig.asp=0.8}
ggplot(top_20_plot_df, aes(
    gsub("^20", "'", as.character(cr_year)),
    value / all_total,
    group = type,
    fill = type
  )) +
  geom_area(stat = "identity") +
  scale_fill_manual(
    "Hybrid OA",
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  facet_wrap( ~ country_name_fct,
              nrow = 3,
              labeller = as_labeller(country_facet_labels)) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1L),
    expand = expansion(mult = c(0, 0.05)),
    breaks = scales::breaks_extended(4)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = "2018 - 2022") +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
    axis.text.x = element_blank(),
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10),
    strip.text = ggtext::element_markdown(size = 6),
    plot.subtitle = ggtext::element_markdown(hjust = 0.5),
    panel.spacing = unit(1, "lines")
  )
```

@@

country_scatter_22 <- country_scatter_df |>
  filter(cr_year == "2022", !is.na(country_code)) |>
  ggplot(aes(all, oa)) +
  geom_point(aes(color = cat, size = ta_oa), alpha = 0.8) +
  scale_color_manual(NULL, values = c("OECD" = "#3fc1c9", "BRICS" = "#fc5185", "Other" = "#e1e1e1")) +
  geom_smooth(aes(all, oa), method=lm) +
  scale_size("TA OA", guide = "none") +
  ggrepel::geom_text_repel(
    data = filter(country_scatter_df, cr_year == 2022, cat %in% c("OECD", "BRICS"), all > 50000),
    aes(x = all, y = oa, label = country_code),
    force = 50,
    box.padding = 0.5,
    point.padding = 0.8,
    size = 3,
    max.iter = 1000000,
    max.overlaps = Inf,
    min.segment.length = 0,
    arrow = arrow(length = unit(0.010, "npc"))) +
  scale_x_continuous(labels =  scales::number_format(big.mark = ","), guide = guide_axis(angle = 45)) +
  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
  theme_minimal() +
  labs(x = "Total articles", y = "OA articles in hybrid journals") +
  theme(legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal")

@@

country_bar <- ggplot(oecd_stats_df, aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
  geom_area() +
  facet_wrap(~ cat) +
  scale_fill_manual(NULL, values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "grey50"),
                    labels = c(oa_other = "OA Other", ta_oa = "Agreement", closed = "Closed")) +
  theme_minimal() +
  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
  theme_minimal() +
  labs(x = "Publication year", y = "Total articles") +
  theme(legend.position = "top",
        legend.justification = "left",
        legend.direction = "horizontal")

@@

```{r country_patch}
library(patchwork)
wrap_plots(
  A = country_scatter_22,
  B = country_bar,
  design = "AB",
  widths = c(0.35, 0.65)
) &
  theme(
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A') 
```

>>>>

```{r area_oa_over_the_years}
oa_ta_plot_df <- ta_df |>
  select(cr_year, ta_oa, no_ta) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  inner_join(year_all)
  # plot
oa_ta_plot <-
  ggplot(oa_ta_plot_df, aes(cr_year, value / n, fill = type, group = type)) +
  geom_area(color = "#000000",
            linewidth = 0.15) +
  scale_fill_manual(
    NULL,
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    limits = c(0, .15),
    breaks = c(0, .05, .1, .15)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )
```

```{r ta_share_plot}
ta_share_plot <- ta_df |>
  select(cr_year, ta_oa_prop, no_ta_prop) |>
  pivot_longer(cols = c(ta_oa_prop, no_ta_prop), names_to = "type", values_to = "value") |>
  filter(!cr_year %in%  c("2017", "2023"), type == "ta_oa_prop") |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = 1)) +
  geom_bar(stat = "identity", fill = "#56B4E9", color = "#000000", linewidth = 0.15) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25, 0.5),
                     labels = scales::percent) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = "Publication Year")
```

@@

summary_all <- inner_join(ta_df, jn_year_all, by = c("cr_year", "issn_l")) |>
  mutate(prop_ta_oa = ta_oa / n,
         prop_no_ta = no_ta / n) |>
  ungroup() |>
  distinct(cr_year, issn_l, prop_ta_oa, prop_no_ta) |>
  pivot_longer(cols = c(prop_ta_oa, prop_no_ta))
# plot
box_ta_prop <- summary_all |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  # With at least one OA article
  filter(value > 0) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, fill = name)) +
  geom_boxplot(outlier.shape = NA, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  coord_cartesian(ylim = c(0,0.65)) +
  scale_fill_manual(NULL, values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement"), guide = FALSE) +
  theme_minimal() +
  labs(y = NULL, x = "Publication Year")
```

@@

publisher_prop <-
  ggplot(publisher_summary_df, aes(
    gsub("^20", "'", as.character(cr_year)),
    prop,
    fill = type,
    group = type
  )) +
  geom_area(color = "#000000",
            linewidth = 0.15) +
  facet_wrap( ~ publisher, nrow = 1) +
  scale_fill_manual(
    "Hybrid OA",
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

publisher_boxplot <- publisher_plot_df |>
  filter(value > 0) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA,
               show.legend = FALSE,
               varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual(
    "Hybrid OA",
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement")
  ) +
  coord_cartesian(ylim = c(0, 0.65)) +
  facet_wrap( ~ publisher, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())

publisher_ta_prop_plot <- publisher_ta_prop |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity",
           fill = "#56B4E9",
            color = "#000000",
           linewidth = 0.15) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    breaks = c(0, 0.25, 0.5, 0.75),
    limits = c(0, .75),
    labels = scales::percent
  ) +
  facet_wrap( ~ publisher, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )
```

@@

```{r subject_plots}
subject_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  select(cr_year, top_level, contains("frac"),-oa_frac) |>
  pivot_longer(
    cols = c(ta_oa_frac, no_ta_fractions),
    names_to = "type",
    values_to = "value"
  ) |>
  mutate(prop = value / articles_frac) |>
  ggplot(aes(
    gsub("^20", "'", as.character(cr_year)),
    prop,
    fill = type,
    group = type
  )) +
    geom_area(color = "#000000",
            linewidth = 0.15) +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_fill_manual(
    "Hybrid OA",
    values = c(
      no_ta_fractions = "#b3b3b3a0",
      ta_oa_frac = "#56B4E9"
    ),
    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

subject_ta_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  mutate(prop = ta_oa_frac / oa_frac) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity", fill = "#56B4E9", color = "#000000",
            linewidth = 0.15) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
                     labels = scales::percent) +
  facet_wrap(~top_level, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank())
subject_box_trend <- subject_df |>
   # with at least one
  select(-ajcs, -subject_area, -esac_publisher) |>
  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(value > 0) |>
  filter(top_level != "Multidisciplinary", !cr_year %in% c("2017", "2023")) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.55)) +
  facet_wrap(~top_level, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )
```

@@

# Prepare facet heading
facet_labels <- country_multiple_df |>
  group_by(country_name_fct) |>
  summarise(total = sum(all_total)) |>
  mutate(prop = total / sum(total)) |>
  ungroup() |>
  mutate(total_string = case_when(
    total < 1000 ~ as.character(total),
    total < 1000000 ~ paste0(format(round(total / 1e3, 1), trim = TRUE), "K"),
    total >= 1000000 ~ paste0(format(round(total / 1e6, 2), trim = TRUE), "M")
  )) |>
  mutate(
    facet_labels = glue::glue(
      '**{country_name_fct}**<br>
      {total_string} | {paste(round(prop * 100, 1), "%")}'
    )
  ) |>
  distinct(country_name_fct, facet_labels)

@@

```{r, country_plot, fig.asp=0.8}
ggplot(top_20_plot_df, aes(
    gsub("^20", "'", as.character(cr_year)),
    value / all_total,
    group = type,
    fill = type
  )) +
  geom_area(color = "#000000",
            linewidth = 0.15,
            stat = "identity") +
scale_fill_manual(
    "Hybrid OA",
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  facet_wrap( ~ country_name_fct,
              nrow = 3,
              labeller = as_labeller(country_facet_labels)) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1L),
    expand = expansion(mult = c(0, 0.05)),
    breaks = scales::breaks_extended(4)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = "2018 - 2022") +
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
    axis.text.x = element_blank(),
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10),
    strip.text = ggtext::element_markdown(size = 5.5),
    plot.subtitle = ggtext::element_markdown(hjust = 0.5),
    panel.spacing = unit(1, "lines")
  )
```

@@

```{r country_bar_chart_top_20}
top_20_df <- country_scatter_df |>
  filter(cr_year == "2022") |>
  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
  group_by(country_name_fct, country_code, cat) |>
  summarise(all = sum(all),
            oa = sum(oa),
            ta_oa = sum(ta_oa), .groups = "drop") |>
  head(20) |>
  mutate(cat_labels = case_when(
   cat == "OECD" ~ glue::glue("<span style = 'color: #3fc1c9'>{country_code}</span>"),
   cat == "BRICS" ~ glue::glue("<span style = 'color: #fc5185'>{country_code}</span>"),
   cat == "Other" ~ glue::glue("<span style = 'color: #e1e1e1'>{country_code}</span>")
  )) |>
  mutate(cat_labels = fct_inorder(cat_labels))

top_20_plot <- top_20_df |>
  mutate(no_ta = oa - ta_oa) |>
  pivot_longer(
  cols = c(no_ta, ta_oa),
  values_to = "value",
  names_to = "type") |>
  ggplot(aes(reorder(as.character(country_code),  value), value / sum(value), fill = type)) +
    geom_bar(stat = "identity", color = "#000000", linewidth = 0.15) +
   scale_fill_manual(
    NULL,
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = FALSE) +
  coord_flip() +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1L),
    expand = expansion(mult = c(0, 0.05)),
    breaks = scales::breaks_extended(4)
  ) +
  theme_minimal() +
#  theme(axis.text.y = ggtext::element_markdown()) +
  labs(x = NULL, y = "% of Hybrid OA published\nin 2022")

```{r}
country_scatter_22 <- country_scatter_df |>
  filter(cr_year == "2022", !is.na(country_code), all > 10000) |>
  ggplot(aes(all, oa)) +
  geom_point(aes(fill = cat, size = ta_oa), pch = 21, alpha = 0.7) +
  scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
  geom_smooth(aes(all, oa), method=lm) +
  scale_size("Agreement", guide = FALSE) +
  ggrepel::geom_text_repel(
    data = filter(country_scatter_df, cr_year == 2022, ta_oa > 5000 | country_code %in% c("CN", "IN", "BR", "RU", "IR")),
    aes(x = all, y = oa, label = country_code),
            min.segment.length = .15,
    force = 5,
    segment.color = "grey10",
    segment.size = 0.5,
    segment.alpha = 0.8,
    color = "grey10", 
    size = 3,
    box.padding = 0.3) +
  scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
  scale_x_log10(labels =  scales::number_format(big.mark = ",")) +
  theme_minimal() +
  labs(x = "Total articles in hybrid journals in 2022", y = "OA articles")

@@

country_bar <- ggplot(oecd_stats_df, aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
  geom_area(color = "#000000",
            linewidth = 0.15) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
  facet_wrap(~ cat) +
  scale_fill_manual("Hybrid OA", values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
                    labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  labs(x = "Publication year", y = "Total articles")

@@

```{r country_patch, fig.asp=0.9}
library(patchwork)
wrap_plots(
  A = country_bar,
  B = country_scatter_22,
  C = top_20_plot,
  design = "
  AC
  BC",
 # guides = "collect",
  widths = c(0.7, 0.3),
  heights = c(0.5,0.5)
) &
  theme(
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10),
    legend.position = "top", 
    legend.justification = "right") &
    plot_annotation(tag_levels = 'A') 
```
